---
title: "Homework 2"
author: "James Clark"
date: "2/3/2022"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(textreadr)
library(tidyverse)
library(readxl)
library(writexl)
library(lubridate)
library(R6)
library(testthat)
library(shiny)
setwd("C:/Users/James/OneDrive/School/MATH_3190")
```


## Load In College Basketball Data {.tabset}


### scan

```{r}
basketball <- scan("http://kenpom.com/cbbga22.txt", what = "list")
##reads as a list

```


### read_html

```{r}
basketball <- read_html("http://kenpom.com/cbbga22.txt")
##returns a character vector 

```


### readr

```{r}
file <- "http://kenpom.com/cbbga22.txt"
download.file(file,"C:/Users/James/OneDrive/School/MATH_3190/cbbga22.txt" )
basketball <- read_fwf(file = "cbbga22.txt", show_col_types = F) 
##reads a tibble

```

### readxl

```{r}
write_xlsx(basketball, "cbbga22.xlsx")
basketball <- read_excel("C:/Users/James/OneDrive/School/MATH_3190/cbbga22.xlsx")
##reads a tibble from an excel spreadsheet

```



## Different Classes in R

### S3
S3 is the simplest system. In this system methods do not belong to the class. Most predifined classes are this type. This would be useful for creating a simple list.

```{r}
x <- list(course = "Math 3190", time = "8:00", student = "James")
```

### S4
S4 provides more structure to its objects.This class includes slots. For example data types can be assigned to elements in a list and only those types will be accepted. This can be useful to help eliminate data input errors.

```{r}
setClass("Course", representation(name = "character", number = "numeric"))

## data now must match given types
##for example: using "Math 3190" for number will return and error
ds <- new("Course", name = "Fundamental of Data Science", number = 3190)
```

### R6
In this class methods belong to objects. Objects are also mutable, so they can be changed in place. This can include built in functions to manipulate data.

```{r}
example <- R6Class("example",list(
        total = 0,
        add = function(x=1){
        self$total <- self$total +1
        }))
z <- example$new()
z$add(7)
z$total
```


## Manipulating the data

### Add column names 

```{r}
basketball <- basketball %>%
  rename( date = X1, away = X2, away_score = X3, home = X4, home_score = X5 ) 
```

### Add column for point differential
```{r}
basketball <- basketball %>%
  mutate(point_diff = home_score - away_score)
```

### sort the data by the location the game was played
```{r}
basketball <- basketball %>%
  arrange(home)
```

### remove columns 6 and 7 
```{r}
basketball <- basketball %>%
  select(-c(X6, X7))
```

### filter for games that were played in 2022 only
```{r}
basketball$date <- mdy(basketball$date)

basketball_2022 <- basketball %>%
  filter(year(date) >= 2022)
```

### Save datasets

```{r}
write.csv(basketball,"C:/Users/James/OneDrive/School/MATH_3190/github/hw2package/inst/cbb.csv", row.names = F)
write.csv(basketball_2022, "C:/Users/James/OneDrive/School/MATH_3190/github/hw2package/inst/cbb22.csv", row.names = F)

```


## Write a function that will filter to only games played by a given team
```{r}
team_select <- function(team, data, chr = FALSE){
  if(chr == TRUE){
    data %>%
      filter(home == team | away == team) %>%
      arrange(date) %>%
      ##this preserves date format for shiny interface
      mutate(date = as.character(date))
  }
  else
    data %>%
    filter(home == team | away == team) %>%
    arrange(date) 
}
team_select("Southern Utah", basketball_2022)
```

## Find a team's winning percentage
```{r}
win_per <- function(team, data){
  team_data <- team_select(team, data)
  team_data %>%
    mutate(win = ifelse(
      (away == team & point_diff <0 ) | (home == team & point_diff > 0),
      1,0
    )) %>%
    summarise(win_per = sum(win)/ n(), wins = sum(win), losses = n() - wins)
}
win_per("Southern Utah", basketball_2022)
```

## Create a package for basketabll functions

```{r, echo=T, eval=F}

setwd("C:/Users/James/OneDrive/School/MATH_3190/github")
devtools::create("hw2package")
usethis::use_mit_license()
devtools::install("hw2package")
library(hw2package)
```


```{r}
##test functions from package
suu_test <- team_select("Southern Utah", basketball_2022)
win_test <- win_per("Southern Utah", basketball_2022)
```


### Unit tests
```{r}
##unit tests for team select
test_that("result",{
  expect_equal(team_select("Southern Utah", basketball_2022),
     suu_test          
) 
})

test_that("object",{
  expect_true(is_tibble(team_select("Southern Utah", basketball_2022)),
) 
})


```

```{r}
##unit tests for win per
test_that("result",{
  expect_equal(win_per("Southern Utah", basketball_2022),
     win_test          
) 
})

test_that("object",{
  expect_true(is_tibble(win_per("Southern Utah", basketball_2022)),
) 
})


```


## Create a graph using basketball data

```{r}

game_graph <- function(team, data){

data <- team_select(team, data)

##change point differential to perspective of selected team instead of home team
data$point_diff <- ifelse(data$away == team, -data$point_diff, data$point_diff)

##add opponent variable
data$opp <- ifelse(data$away == team, data$home, data$away)

##create variable for if they won
data$Result <- as.factor(ifelse(data$point_diff > 0, "Win", "Loss"))

data <- data %>%
  arrange(date)

ggplot(data,aes(x = point_diff, y = opp, fill= Result) ) + geom_bar(stat = 'identity') + geom_vline(xintercept = 0) + 
  labs(x = "Point Differential",y = "Opponent", title = paste(team,"Point Difference by Game") )

}

suu_graph <- game_graph("Southern Utah", basketball_2022)
```

### Unit tests for game_graph

```{r}
setwd("C:/Users/James/OneDrive/School/MATH_3190/github")
devtools::install("hw2package")
library(hw2package)

test_that("result",{
  expect_equal(game_graph("Southern Utah", basketball_2022),
     suu_graph         
) 
})

```

## Shiny Interface

```{r}

##This will require the basketball_2022 df that was created above
setwd("C:/Users/James/OneDrive/School/MATH_3190/github/hw2package/inst")
source("ui.R")
source("server.R")

shinyApp(server = server, ui = ui)

##similarly
shiny_app("C:/Users/James/OneDrive/School/MATH_3190/github/hw2package/inst")
```





## Kmeans tutorial

### Step 00: Template
This creates a blank template with a sidebar panel. 

```{r, template, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      
    )
  )
)

server <- function(input, output, session) {

}

shinyApp(ui = ui, server = server)
```

### Step 01: Title Panel

This step adds a "k-means clustering" title. 

```{r, add title panel, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      
    )
  )
)

server <- function(input, output, session) {
  
}

shinyApp(ui = ui, server = server)
```

### Step 02: UI inputs

This step adds inputs that are used for clustering. Namely, it adds an input for a X and Y variable, as well as the number of clusters. 

```{r, add ui inputs, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      
    )
  )
)

server <- function(input, output, session) {
  
}

shinyApp(ui = ui, server = server)
```

### Step 03: UI outputs

While this step does not appear to do much within the app it "saves a spot" for a plot to be put based on the inputs. 

```{r, add ui outputs, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {
  
}

shinyApp(ui = ui, server = server)
```

### Step 04: Output plot

This step adds a plot of the irises based on their length and width. However, these are not reactive.

```{r, add output plot, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {
  output$plot1 <- renderPlot({
    plot(iris[,c("Sepal.Length","Sepal.Width")])
  })
}

shinyApp(ui = ui, server = server)
```

### Step 05: Reactive Axes

In this step the plot is now reactive, so it responds to input variables we created earlier.

```{r, add reactive axes, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {
  output$plot1 <- renderPlot({
    plot(iris[,c(input$xcol,input$ycol)])
  })
}

shinyApp(ui = ui, server = server)
```

### Step 06: K means

This step adds the kmeans algorithm, so now our data is divided into how many clusters our input dictates. 

```{r, add K means clustering, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {

  output$plot1 <- renderPlot({
    plot(iris[,c(input$xcol,input$ycol)],
         col = kmeans(iris[,c(input$xcol,input$ycol)], input$clusters)$cluster)
  })
}

shinyApp(ui = ui, server = server)
```

### Step 07: Cluster centers

This step add a big X in the center of each cluster.

```{r, add cluster centers, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {
  
  output$plot1 <- renderPlot({
    plot(iris[,c(input$xcol,input$ycol)],
         col = kmeans(iris[,c(input$xcol,input$ycol)], input$clusters)$cluster)
    points(kmeans(iris[,c(input$xcol,input$ycol)], input$clusters)$centers, pch = 4, cex = 4, lwd = 4)
  })
}

shinyApp(ui = ui, server = server)
```

### Step 08: Save Kmeans

This adds a reactive function which saves the clusters. 

```{r, save k means result as reactive, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {
  
  clusters <- reactive({
    kmeans(iris[,c(input$xcol,input$ycol)], input$clusters)
  })
  
  output$plot1 <- renderPlot({
    plot(iris[,c(input$xcol,input$ycol)],
         col = clusters()$cluster)
    points(clusters()$centers, pch = 4, cex = 4, lwd = 4)
  })
}

shinyApp(ui = ui, server = server)
```

### Step 09: Final App

This final step fills in the points as well as changes the color scheme to make it a little more presentable. 

```{r, final app, eval=F, echo=T}
ui <- fluidPage(
  
  # Put a titlePanel here
  titlePanel("k-means clustering"),
  
  sidebarLayout(
    # Sidebar. Put your inputs inside the sidebarPanel
    sidebarPanel(
      selectInput('xcol', 'X Variable', names(iris)),
      selectInput('ycol', 'Y Variable', names(iris),
                  selected=names(iris)[[2]]),
      numericInput('clusters', 'Cluster count', 3,
                   min = 1, max = 9)
    ),
    
    # Main panel. put your output plot here
    mainPanel(
      plotOutput('plot1')
    )
  )
)

server <- function(input, output, session) {
  
  selectedData <- reactive({
    iris[, c(input$xcol, input$ycol)]
  })
  
  clusters <- reactive({
    kmeans(selectedData(), input$clusters)
  })
  
  output$plot1 <- renderPlot({
    palette(c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3",
              "#FF7F00", "#FFFF33", "#A65628", "#F781BF", "#999999"))
    
    par(mar = c(5.1, 4.1, 0, 1))
    plot(selectedData(),
         col = clusters()$cluster,
         pch = 20, cex = 3)
    points(clusters()$centers, pch = 4, cex = 4, lwd = 4)
  })
}

shinyApp(ui = ui, server = server)
```

## Create Kmeans function for Kmeans package

```{r}

##Function to create clusters
kmeans <- function(x, y, clust, n){
  
  ##Set initial midpoints
    x_clust <- runif(clust, min = min(x), max = max(x))
    y_clust <- runif(clust, min = min(y), max = max(y))
  
    clusters <- data.frame(x_clust, y_clust)
  
  ##find the nearest cluster midpoint for each point
    data <- data.frame(x, y)
    
    ##initialize variables
    
      ##set as large value that no point could have a greater distance than
      data$min_dist <- 2 * (max(x)^2+max(y)^2)
      data$dist <- NA
      data$clust <- NA
    
    ##Itterate n many times
  for (j in 1:n){    
      
    ## Finds nearest midpoint for each point  
    for (i in 1:clust){
      ##Calculates distance 
      data$dist <- sqrt( (x-clusters$x_clust[i])^2+(y-clusters$y_clust[i])^2)
      
      ##redefine cluster if midpoint is closer    
      data$clust <- ifelse(data$dist < data$min_dist, i, data$clust)
    
      ##adjust min dist if needed
      data$min_dist <- ifelse(data$dist < data$min_dist, data$dist, data$min_dist)
     
      
    }
    
    ##Re-assign cluster centers
    clusters_new <- data %>%
  group_by(clust) %>%
  summarise(x_avg = mean(x), 
            y_avg = mean(y)) %>%
  arrange(clust)

clusters$x_clust <- clusters_new$x_avg
clusters$y_clust <- clusters_new$y_avg
    
  }
      data %>%
        select(x, y, min_dist, clust)
}


```

## Create function to plot output for kmeans

```{r}

##Function to plot output from kmeans function
k_viz <- function(data, xlab, ylab){
  ggplot(data, aes(x = data[,1], y = data[,2], color = as.factor(data[,4]) ) ) + geom_point() + 
   scale_color_discrete(name = "Cluster") + labs(x = xlab, 
                                                 y = ylab)
}
```

## Create a K-means package

```{r, echo=T, eval=F}

setwd("C:/Users/James/OneDrive/School/MATH_3190/github")
devtools::create("hw2kmeans")
usethis::use_mit_license()
setwd("C:/Users/James/OneDrive/School/MATH_3190/github/hw2kmeans")
devtools::document()
```

## Unit tests for kmeans package

```{r}
setwd("C:/Users/James/OneDrive/School/MATH_3190/github")
devtools::install("hw2kmeans")
library(hw2kmeans)
```

```{r}
data(iris)


test <- kmeans(iris$Sepal.Length, iris$Sepal.Width, 4, 10)

##try to pass a dataframe as an argument
test_that("result",
  expect_error( kmeans(iris$Sepal.Length, iris, 4, 10)
) 
)

test_that("object",{
  expect_true(is.data.frame(kmeans(iris$Sepal.Length, iris$Sepal.Width, 4, 10)),
) 
})


```

```{r}
##unit test for k_viz
k_test <- k_viz(test, "X Label", "Y Label")

test_that("result",{
  expect_equal(k_viz(test,"X Label", "Y Label" ),
     k_test) })

test_that("object", {
  expect_true(is.list(k_viz(test,"X Label", "Y Label" )))
})

```

##Kmeans shiny interface


```{r}
data(iris)
setwd("C:/Users/James/OneDrive/School/MATH_3190/github/hw2kmeans/inst")
source("ui.R")
source("server.R")

shinyApp(server = server, ui = ui)

shiny_app("C:/Users/James/OneDrive/School/MATH_3190/github/hw2kmeans/inst")
```